\documentclass[11pt]{article}

\title{\textbf{Business Plan: Web Marketists Inc.}}
\author{Sergio Arroutbi Braojos}
\date{\today}
\usepackage{listings}
\usepackage{hyperref}
\addtolength{\voffset}{-50pt}

\begin{document}

\hypersetup
{   
pdfborder={0 0 0}
}

\maketitle

\tableofcontents

\section{Introduction}
\textbf{Web Marketists Inc.} is a company whose main vision is to sell services associated to Market Research in on-line sales companies.\\
\\
In order to achieve its mission, Web Marketists will be based on \textbf{WebMarketizer FLOSS project}, which is a tool, to be developed, that will allow, once correctly configured, connection to on-line sales web pages of a certain environment, and automatic extraction of sale data, such as the type of product, the trademark of the product, the link to reach the product on the original store and of course, the price.\\
\\
Web Marketists Inc. will develop as well web pages   associated to different products, such as electronic devices, informatics, bikes, cars, motorbikes, etc. As long as the different \textbf{environments} are integrated and selling data is collected, a wider bunch of customers will be available for the data to be sold.\\
\\
The company \textbf{will sell different services associated to collected data} to companies interested on it, in order for them to understand the products price in other on-line sale companies, the mean price in the market for a specific product or type of product, the number of different product of each store and, in general, all the on-line market share information that can help a particular company to adjust its strategy to the customer requirements.\\
\\
Apart from that, Web Marketists Inc. will create a web page for all the people to be able to consult, in a costless way, data and price for a product on the different on-line sale webs associated to a certain product. \textbf{Data searched by users is also an exploitable source of data}, to check which are the most preferred consults by users, and sell them as well to the potential on-line sales customers.
\section{Clients segments}
The clients segments where Web Marketists Inc. product is focused is, basically, on-line sales companies. It will focus first on Spanish on-sale companies, planning to extend its products to English spoken countries, such as USA, United Kingdom or Australia.\\
\\
As Web Marketists Inc. develops configuration to enable new data extraction via WebMarketizer FLOSS projects, a bigger set of customers will be available for Web Marketists Inc. Web Marketists Inc. \textbf{will study the different markets} and will poll the different on-line sales companies of each market to adapt the configuration and development of WebMarketizer FLOSS projects to the environment that is supposed to be most favorable to later sell the market research reports. 
\section{Value proposal}
The main vision of Web Marketists Inc. is to \textbf{sell to its customers the best market research services}, in the shortest time, with the most extended information. The knowledge of the WebMarketizer FLOSS tool, together with a deep analysis of the customer requirements, will help on producing the most value added market research reports to the customers.
\section{Channels}
The Channels for Web Marketists Inc. to offer its products to potential customers are \textbf{very oriented to the potential customers}. Web Marketists Inc. will identify the potential customers, and will focus on contacting its customer and sell its services. For each sales environment, it is a key to identify the top ten to twenty on-line selling stores and contact each of them and their chief executives in a personally way to demonstrate how helpful can be the products sold by Web Marketists Inc.\\
\\
The company will also focus on the analysis of the different environments it generates the reporting and analysis data, in order \textbf{to focus on the different on-line sale store environment}. If, for example, Web Marketists develops and collects data of on-line store selling bicycles in Spain, Web Marketists Inc. will participate in Expobike and in Festibike, to make both its services as well as its web pages to be known.
\section{Key resources}
Key resources for Web Marketists Inc. are:
\begin{itemize}
\item{\textbf{Employees knowledge}}. A high knowledge is needed to be able to adequate configurations and fix WebMarketizer FLOSS project to the environment being generated, and being able to collect, generate, analyze and graphically represent market research data to the customers.
\item{\textbf{Customers requirements}}. Customer requirements are very important for Market Web Inc. Reports generated are the better as more helpful result for the customers and potential customers.
\item{\textbf{WebMarketizer FLOSS project}}. The WebMarketizer FLOSS projects is an automatation tool that, once configure, can download data from on-line sale stores and generate different output for this data, from files ready to be inserted in a Database to text files, as well as other standard data files such as XML or web.
\item{\textbf{Market Research data}}. The data collected from the different on-line stores is a key resource data for Web Marketists Inc., as this data will be the base to generate the different reports.
\item{\textbf{Generated web pages}}. On the one hand, a specific web page around a certain environment will be part of the trademark and corporate image of Web Marketists Inc.\\
On the other hand, all the web pages will help users to consult prices and types of products in the different environments, while the web pages will also mean a way to show the amount of data the company has acquired and manages on a certain environment.\\
Apart from that, the web pages will also mean revenue streams, in two different ways:
\begin{enumerate}
\item{\textbf{Web page advertisement}}. As on-sale store companies will also be able to advertise on them.
\item{\textbf{User preferred searches}}. To provide customers the most frequently searched products by users, and, on this manner, help them to investigate user preferences.
\end{enumerate}
\end{itemize}
\section{Cost structure}
The cost structure on Web Marketists Inc. is basically oriented to the employees working on them, specially on the first year. First year will be used to develop the FLOSS WebMarketizer project. The second year onwards, complete task force will be hired, and services will start to be sold.\\
\\
\subsection{First year}
In the first 12 months, one software engineer will be hired in order to develop WebMarketizer FLOSS tool. So, in this 12 months, the estimated cost will be around 40000 euros (including taxation and social insurance).\\
\\
Additional costs will be the amount to start the company (3000 euros), a desktop or laptop and Internet access for the programmer (2000 euros). No office will be needed for this year. To summarize, the costs, altogether, will sum up to \textbf{45000 euros in the first year}.
\subsection{Second year onwards}
After the first year, the FLOSS tool must have been developed for the basic functionality, and configuration must be generated to provide at least the collected data in one environment. A provisional web page will also be developed to measure results of the FLOSS tool, as well as to measure penetration in the market and customers opinion.\\
\\
The second year and onwards, two additional people will be hired for the company:
\begin{itemize}
\item{\textbf{A commercial}}: in order to introduce, advertise, promote and make the services known to the potential customers. (30000 euros + 10\% of the sales)
\item{\textbf{Another programmer}}: more focused on web programming, in charge of development and maintenance of web pages, on the one hand, but that also to help on configurations and customization of WebMarketizer tool to collect the data (40000 euros a year)
\end{itemize}
Apart from people hired, other direct costs will be:
\begin{itemize}
\item{\textbf{Business Office}} (12000 euros a year)
\item{\textbf{Hardware + hosting + mobile phone line + Internet access}} (3000 + 1000 = 5000 in euros a year)
\item{\textbf{Advertisement/marketing campaigns}} (3000 euros in a year)
\end{itemize}
So, from the second year and ahead, the total cost per year will be:
\begin{itemize}
\item{Employees}: 110000 euros
\item{Other costs}: 20000 euros
\end{itemize}
\textbf{Total: 130000 euros a year.}
\section{Revenues streams}
Different ways of incomes will be available based on the different customer requirements. The main revenues streams will be with market research reports, but additional incomes will be available to the company due to different products and services to be acquired by potential customers. They are described in the next sections:
\subsection{Market Research Reports}
Main revenues streams will come from generation of Market Research reports for on-line sales companies. The information will be received, by the companies, in different manners by different service subscriptions:
\begin{itemize}
\item{Reception of a copy once a month of the Database with the analyzed data}. \textbf{Price: 600 euros a year / 60 euros a month}. 
\item{Reception of a complete report with different market research information}. It also includes a copy of the Database. Information should include, among other:
\begin{itemize}
\item{Offer catalog comparison against competitors and competitors means}.
\item{Offer catalog comparison, sorted by type of product, against competitors and competitors means}.
\item{Best price and best price by product and type of product among all competitors}.
\item{Mean price and mean price by type of product, against competitors price and mean price}.
\end{itemize}
\textbf{Price: 1200 euros a year / 120 euros a month}. 
\end{itemize}
\subsection{User preferred searches}
Reception, by a customer, of the top searches made by users in a month. \textbf{Price: 100 euros a month}.
\subsection{Advertisement}
With different prices, a customer can insert their trademark and/or offers in the consult web pages developed by Web Marketists. There will be three different plans:
\begin{itemize}
\item{Normal banner}. \textbf{Price: 50 euros a month}.
\item{Special banner}. Placed on a much better place than a normal banner. \textbf{Price: 100 euros a month}.
\item{Dedicated banner}. Placed on the best place of the web page. \textbf{Price: 300 euros a month}.
\item{Special offer}. Where companies can insert ``the week special offer'' on their on-line prices. \textbf{Price: 100 euros a week}.
\end{itemize}
\subsection{Training and installation of WebMarketer FLOSS tool}
If a company prefers installation on their informatics facilities of the WebMarketizer FLOSS tool, in order to acquire the data in an automatic way, a package including installation and training can be acquired. It includes installation of the tool, a one week training to demonstrate the tool functionality, and a first acquisition of data and generation of reports.\\
\textbf{Price: 20000 euros}.
\subsection{Platinum exclusive contract}
A company \textbf{of a certain environment}, will be able to acquire all the services (Database, Market Research Reports, User preferences, WebMarketizer FLOSS training and installation and exclusive web Advertisement) in a \textbf{exclusive contract}. This kind of exclusive contract will only be possible in case a certain environment is still not being economically exploited by Web Marketists Inc. and no contracts have been signed to any other customer of that particular environment.\\
\textbf{Price: 200000 euros a year}.
\section{Customer relationships}
The customer relationships will depend on the services provided to the customers. 
\textbf{The main customer relationship will be associated with the Platinum exclusive customer} (if any). The commercial employee will have a weekly meeting to sort out all the requirements that are to be covered in the contract. Meanwhile, for the rest of customers of a particular environment, no special relationships will be hold, apart from the contractual ones, such as sending the monthly market research report, sending report of the most searched products on the web or placing the banner on the web page.
\section{Key activities}
Key activities have already been described above, but a summary will be exposed below:
\begin{itemize}
\item{Development of WebMarketizer FLOSS project.}
\item{Environment analysis and market study to select more profitable environments.}
\item{Configuration of WebMarketizer, Data Acquisition and Report Generation for a particular environment.}
\item{Sell force empowering.}
\item{Task force and knowledge preservation.}
\end{itemize}
\section{Key partners}
No special partners are needed on this kind of business, as Web Marketists services will be directly sold by the company, and provision of the different services will be also held by the company.
\section{Conclusion}
\textbf{E-commerce grew in Spain 20 \% in 2012, billing up to 12000 million of euros}. It is a very remarkable data, taking into account the actual economic situation of the country. Identifying those environments where more billing is performed will also help on identifying potential customers willing to pay for services that allow them to investigate further on the market share of a certain set of products.\\
\\
Web Marketists Inc., via WebMarketer FLOSS tool, \textbf{will provide the functionality, the data and the services for on-line sale businesses to acquire a better knowledge of their particular markets}, as well as other additional services such as knowledge of the users preferences when searching for a particular product, as well as providing an additional advertising mechanism on the Internet.\\
\\
Web Marketists Inc., the Enterprise to help improving e-commerce companies sales force, is willing to be born, to be funded and to keep a dominant role on on-line Market Research service providing. Who joins?
\end{document}
